---
title: "tidy_data"
output: html_document
date: "Last update: `r Sys.Date()`"
---

# libs

```{r}
#| label: setup
#| warning: false
#| message: false
library("dplyr")
library("tidyr")
library("readr")
library("here")
library("fs")
library("forcats")
library("stringr")
library("purrr")
library("readxl")
library("ggplot2")
library("lme4")
library("marginaleffects")

```

# Load and tidy data

## Local data

```{r}
#| label: load-local-data
#| message: false
# Load local csvs 
# except 
#  - 028_exp_g2_2023-05-05_15h19.23.523.csv (because it has more columns)

local_temp1 <- dir_ls(here("dat", "raw_local"), regexp = ".csv") |> 
  as_tibble() |> 
  filter(
    value != here("dat", "raw_local", "028_exp_g2_2023-05-05_15h19.23.523.csv")
  ) |> 
  pull() |> 
  map_dfr(
    read_csv, 
    id = "source", 
    col_types = cols(.default = "c")
  ) |> 
  filter(exp_num == 1, lang %in% c("en", "es")) |> 
  transmute(
    participant, 
    group = expName, 
    list, cond, lang, type, num, 
    exp_num, 
    cs, 
    correct_key, 
    response = exp1_key.keys, 
    rt = as.numeric(exp1_key.rt), 
    is_correct = as.numeric(exp1_key.corr)
  )

# load 
#  - 028_exp_g2_2023-05-05_15h19.23.523.csv (because it has more columns)
local_temp28 <- read_csv(
  here("dat", "raw_local", "028_exp_g2_2023-05-05_15h19.23.523.csv")
  ) |>
  filter(exp_num == 1, lang %in% c("en", "es")) |> 
  transmute(
    participant = as.character(participant), 
    group = expName, 
    list, cond, lang, type, num, 
    exp_num = as.character(exp_num), 
    cs = as.character(cs), 
    correct_key, 
    response = exp1_key.keys, 
    rt = as.numeric(str_remove_all(exp1_key.rt, "\\[|\\]")), 
    is_correct = as.numeric(exp1_key.corr)
  )

# Combine local in single df
df_local <- bind_rows(local_temp1, local_temp28) |> 
  mutate(source = "local")
```


## Online data

```{r}
#| label: load-online-data
#| message: false
# Load online csvs 
# These files all have different columns names/numbers
#  - 021_exp_g1_2023-09-05_18h32.55.031
#  - 027_exp_g2_2023-10-18_14h15.06.123
#  - 029_exp_g2_2023-10-21_17h02.18.335
#  - 033_exp_g2_2023-10-12_11h56.31.376
#  - 042_exp_g1_2023-09-29_14h14.23.677
#  - 054_exp_g2_2023-10-12_12h40.24.192
#  - 062_exp_g2_2023-11-16_11h11.01.336
#  - 064_exp_g2_2023-10-23_17h51.37.190
#  - 067_exp_g2_2023-10-18_16h58.16.959
#  - 070_exp_g1_2023-09-22_13h39.35.259
#  - 073_exp_g2_2023-09-11_17h01.16.888
#  - 074_exp_g2_2023-10-06_10h15.54.848
#  - 075_exp_g2_2023-10-12_11h13.11.698
#  - 086_exp_g1_2023-09-19_11h05.55.100
#  - 091_exp_g2_2023-09-14_12h03.00.504

# Put them in a list: 
temp_list <- dir_ls(here("dat", "raw_online"), regexp = ".csv") |> 
  as.list() |> 
  map(read_csv)

# Make vector of only the columns we want
my_cols <- c(
  "participant", "expName", "list", "cond", "lang", "type", "num", "exp_num",
  "cs", "correct_key", "exp1_key.keys", "exp1_key.rt", "exp1_key.corr"
)

# Select only the columns in my_cols in each df in the list
temp_list <- Map(function(x){x[,names(x) %in% my_cols]}, temp_list)

# Bind them all into one df
df_online <- do.call("rbind", temp_list) |> 
  filter(exp_num == 1, lang %in% c("en", "es"), !is.na(list)) |> 
  transmute(
    participant, 
    group = expName, 
    list, cond, lang, type, num, 
    exp_num = as.character(exp_num), 
    cs = as.character(cs), 
    correct_key, 
    response = exp1_key.keys, 
    rt = exp1_key.rt,
    is_correct = exp1_key.corr
  ) |> 
  mutate(source = "local")
```

## Combine everything

```{r}
#| label: combine-online-local
#| message: false
# Combine df_local and df_online
df_complete <- bind_rows(df_local, df_online) |> 
  mutate(
    participant = str_pad(participant, pad = "0", width = 3), 
    num = str_remove(num, "\\.0"), 
    num = str_pad(num, pad = "0", width = 3)
  ) |> 
  # Join with lextale data
  left_join(
    read_csv(here("dat", "lang_bg", "lang_bg.csv")) |> 
    select(participant, lextale) |> 
    mutate(
      participant = as.character(participant), 
      participant = str_pad(participant, side = "left", pad = "0", width = 3), 
      lextale_z = (lextale - mean(lextale)) / sd(lextale)
      ), 
    by = "participant"
  ) |> 
  # Join with matrix duration data
  left_join(
    read_excel(here("exp", "stim_new", "stimuli_all.xlsx")) |> 
    mutate(
      num = str_remove(num, "\\.0"), 
      num = str_pad(num, pad = "0", width = 3)
    ) |> 
    filter(
      mat_dur != "-", 
      lang %in% c("en", "es"), 
      list %in% c("l1", "l2"), 
      cond %in% c("t", "f"), 
      type %in% c("na", "f0c", "f0u")
    ) |> 
    select(num, mat_dur), 
    by = "num"
  ) |> 
  filter(!is.na(mat_dur)) |> 
  mutate(rt = rt - as.numeric(mat_dur)) |> 
  filter(rt >= 0) |> 
  write_csv(here("dat", "exp1_clean_jvc.csv"))

```






# plot

```{r}
#| label: plot-rts
#| dpi: 300
#| fig-asp: 0.56
#| out-width: "100%"

#calculate limit
df_limit<-df_complete %>% 
  group_by(participant) %>% 
  summarize(ave = mean(rt), limit = ave + sd(rt)*2)

df_lted<-df_complete%>% 
  left_join(., df_limit, by = 'participant')
  
df_ltd<-df_lted%>%
  filter (., rt <= limit)

#combine limit and make plots
level_order <- c('na.es', 'na.en', 'f0c.es', 'f0u.es') 
df_ltd%>% 
  group_by(participant) %>% 
  #filter(., rt <= limit) %>%
  filter(is_correct == 1) |> 
  ggplot() +
  aes(x = interaction(type, lang), y = rt) +
  scale_x_discrete(limits = level_order) +
  geom_point(color = "magenta", alpha = 0.3) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange", pch = 21, fill = 'white') +
  stat_summary(
    fun.data = mean_se, size = 4,
    geom = "text", 
    aes(label = sprintf("%0.2f", round(after_stat(y), digits = 2))),
    position = position_nudge(x = 0.25)
  ) +
  labs(
    x = "Conditions", 
    y = "Reaction Time", 
    title = "Experiment 1 RT as a function of Conditions",
    caption = "reaction time as a function of Conditions"
  ) 

df_ltd |> 
  filter(rt <= limit) |>
  filter(is_correct == 1) |> 
  ggplot() + 
  aes(x = lextale_z, y = rt, color = interaction(lang, type)) + 
  geom_point(alpha = 0.4) + 
  geom_smooth(method = lm, se = F, fullrange = T, formula = "y ~ x") + 
  scale_color_viridis_d(name = NULL, end = 0.8) +
  labs(
    x = "Proficiency", 
    y = "Reaction Time", 
    title = "Experiment 1 RT as a function of proficiency",
    caption = "reaction time as a function of proficiency"
  ) 

```

```{r}
#| label: plot-accuracy
#| dpi: 300
#| fig-asp: 0.56
#| out-width: "100%"
df_ltd |> 
  ggplot() + 
  aes(x = type, y = is_correct, fill = lang) + 
  geom_jitter(
    data = df_complete |> 
      group_by(participant, type, lang) |> 
      summarize(avg = mean(is_correct, na.rm = T), .groups = "drop"), 
    aes(x = type, y = avg, color = lang), 
    height = 0, width = 0.2
  ) + 
  stat_summary(
    fun.data = mean_se, geom = "pointrange", pch = 21, 
    position = position_dodge(0.5)
  ) + 
  scale_fill_viridis_d(name = NULL, end = 0.8) + 
  scale_color_viridis_d(name = NULL, end = 0.8)
```



# Models

```{r}
#| label: model

df_mod <- df_ltd |> 
  filter(is_correct == 1) |> 
  unite(col = condition, c("lang", "type"), sep = "-") |> 
  mutate(
    log_rt = log(rt), 
    condition = fct_relevel(condition, "es-na")
  ) 

mod_lt <- lmer(
  log_rt ~ lextale_z + condition + (1 | participant), 
  data = df_mod
)

summary(mod_lt)

```

# Fit models

```{r, 'mod_null'}
# Fit a null model
mod_null <- lmer(log_rt ~ 1+ (1 | participant), data = df_mod)
summary(mod_null)

# Add a continuous predictor

mod_1 <- lmer(log_rt ~ lextale_z + (1 | participant), data = df_mod)
summary(mod_1)
anova(mod_1)
# Add a categorical covariate
mod_2 <- lmer(log_rt ~ lextale_z + condition + (1 | participant), data = df_mod)
summary(mod_2)
anova(mod_2)

# Include an interaction term
mod_3 <- lmer(log_rt ~ lextale_z * condition+ (1 | participant) , data = df_mod)
summary(mod_3)

## Test main effects and interaction term
anova(mod_null, mod_1, mod_2, mod_3)
```

# Results

The reaction time data were logged and analyzed using a linear mixed-effects model. Estimated reaction time was the criterion with language proficiency and conditions predictors, we also included a by-participant intercept. The condition group factor was dummy coded with condition group na.es set as the reference group. Main effects and the language proficiency by condition group interaction were assessed using nested model comparisons. Experiment-wise alpha was set at 0.05.

There was a main effect of condition group (Χ2  = 83.274,  p < 0.001), no main effect of proficiency (X2 = 1.2328, p = 0.267) nor an condition group by proficiency interaction (X2 = 0.925; p = 0.819). The model containing the interaction provided the best fit of the data (R2 = 0.7361). Overall, proficiency score increased as a function of experience. However, the size of the effect was modulated by age group. Specifically, adults showed an increase of score of approximately 0.6476  +/- 0.1334 se (t = 4.856, p < 0.001) per year of experience. Children showed an additional increase of 0.7198  +/- 0.1886 se per year of experience (1.3674 in total , t = 3.817, p < 0.001).
